---
title: "SPECC 2019 Behav"
author: "Michael Hallquist"
date: "9/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=180)
#entropy change betas
library(readr)
library(lme4)
library(tidyverse)
library(psych)
library(emmeans)
library(ggplot2)
library(sjPlot)

base_dir <- "/Users/mnh5174/Box/SCEPTIC_fMRI/specc_betas"
beta_dir <- file.path(base_dir, "sceptic-clock-feedback-v_entropy_change-preconvolve_fse_groupfixed", "v_entropy_change")
bpq <- read.csv("/Users/mnh5174/Data_Analysis/SPECC/self_reports/data/surveys/baseline/scored/BPQ_scored.csv")
behav <- trial_df <- read.csv(file.path(base_dir, "specc_decay_factorize_selective_psequate_fixedparams_ffx_trial_statistics.csv.gz")) %>%
  mutate(
    run_trial=case_when(
      trial >= 1 & trial <= 50 ~ trial,
      trial >= 51 & trial <= 100 ~ trial - 50L, #dplyr/rlang has gotten awfully picky about data types!!
      trial >= 101 & trial <= 150 ~ trial - 100L,
      trial >= 151 & trial <= 200 ~ trial - 150L,
      trial >= 201 & trial <= 250 ~ trial - 200L,
      trial >= 251 & trial <= 300 ~ trial - 250L,
      trial >= 301 & trial <= 350 ~ trial - 300L,
      trial >= 351 & trial <= 400 ~ trial - 350L,
      TRUE ~ NA_integer_),
    v_entropy_no5=if_else(run_trial <= 5, NA_real_, v_entropy),
    d_auc_sqrt=if_else(d_auc > 0, NA_real_, sqrt(-1*d_auc)), #only compute the sqrt of d_auc for negative (i.e., reasonable) observations
    v_entropy_sqrt=sqrt(v_entropy),
    rew_om=if_else(score_vba > 0, 1, 0)
  ) %>% #for win/loss maps
  group_by(id, run) %>%
  dplyr::mutate(   #compute rt_swing within run and subject
    rt_csv=rt_csv/1000, #rescale for ease in lmer
    rt_lag = dplyr::lag(rt_csv, 1, order_by=trial),
    rt_vmax_lag = dplyr::lag(rt_vmax, 1, order_by=trial),
    rt_vmax_change = abs(rt_vmax - rt_vmax_lag),
    omission = factor(as.numeric(score_csv > 0), levels=c(0,1), labels=c("Omission", "Reward")),
    omission_lag = lag(omission, order_by=trial),
    v_entropy_lag = dplyr::lag(v_entropy, 1, order_by=trial),
    v_entropy_change = v_entropy - v_entropy_lag, #change in entropy
    v_entropy_change_pos = v_entropy_change*(v_entropy_change > 0),
    v_entropy_change_neg = abs(v_entropy_change*(v_entropy_change < 0)),
    vmax_lag = dplyr::lag(v_max, 1, order_by=trial),
    pe_lag = dplyr::lag(pe_max, 1, order_by=trial),
    ppe_lag = if_else(pe_lag > 0, pe_lag, 0),
    npe_lag = if_else(pe_lag < 0, abs(pe_lag), 0),
    rt_swing = abs( c(NA, diff(rt_csv))),
    rt_swing_sqrt=sqrt(rt_swing)) %>%
  ungroup() %>% select(-dataset, -rt_next, -score_next, -clock_onset, -isi_onset, -feedback_onset, -iti_onset, -image) %>%
  rename(entropy=v_entropy, score=score_csv) %>%
  mutate(emotion=recode(emotion, scram="Scrambled", fear="Fear", happy="Happy"))

design <- read.table(file.path(beta_dir, "v_entropy_change-Intercept-Age_c-BPD_c-BPD_Age_design.txt"), header=TRUE) %>%
  select(feat_input_id, id, SPECC_ID, BPD, Age, Female, Age_c, Age_c, BPD_Age) %>%
  mutate(group=factor(BPD, levels=c(0,1), labels=c("HC", "BPD")),
         Female=factor(Female, levels=c(0,1), labels=c("Male", "Female")))

behav <- behav %>% inner_join(design)

names(bpq)
bpq$SPECC_ID <- sub("_", "", as.character(bpq$SPECC_ID), fixed=TRUE) %>% dplyr::recode("144DB"="144DH")
behav <- behav %>% left_join(bpq %>% select(SPECC_ID, BPQ_total), by="SPECC_ID")
```

## performance by run/condition

```{r}
behav_by_run <- behav %>% group_by(id, run) %>% summarize(
  earnings = sum(score),
  rewFunc=rewFunc[1],
  emotion=emotion[1],
  BPQ_total=BPQ_total[1],
  group=group[1],
  avg_ev = mean(ev),
  entropy = mean(entropy),
  Age_c = Age_c[1]
) %>% ungroup() %>%
  mutate(BPQ_total = BPQ_total - mean(BPQ_total, na.rm=TRUE))

summary(m1 <- lmer(earnings ~ run + rewFunc*emotion*BPQ_total + (1|id), behav_by_run))
car::Anova(m1, type=3)
plot_model(m1, type="pred", terms=c("BPQ_total", "rewFunc", "emotion"))

summary(m2 <- lmer(earnings ~ run + rewFunc*emotion*BPQ_total + (1|id), behav_by_run %>% filter(rewFunc %in% c("IEV", "DEV"))))
car::Anova(m2, type=3)
plot_model(m2, type="pred", terms=c("BPQ_total", "rewFunc", "emotion"))

summary(m3 <- lmer(earnings ~ run + rewFunc*emotion*group + (1|id), behav_by_run %>% filter(rewFunc %in% c("IEV", "DEV"))))
car::Anova(m3, type=3)
plot_model(m3, type="pred", terms=c("group", "rewFunc", "emotion"))

```

Whether we treat BPD as a dimension or a two-level factor, there is no evidence that earnings differ systematically by group

The earlier analyses hold, with Happy hurting IEV learning, but helping DEV learning.

What about choice quality (mean EV)?
```{r}

summary(m4 <- lmer(avg_ev ~ run + rewFunc*emotion*BPQ_total + (1|id), behav_by_run))
car::Anova(m4, type=3)
plot_model(m4, type="pred", terms=c("BPQ_total", "rewFunc", "emotion"))

summary(m5 <- lmer(avg_ev ~ run + rewFunc*emotion*BPQ_total + (1|id), behav_by_run %>% filter(rewFunc %in% c("IEV", "DEV"))))
car::Anova(m5, type=3)
plot_model(m5, type="pred", terms=c("BPQ_total", "rewFunc", "emotion"))

summary(m6 <- lmer(avg_ev ~ run + rewFunc*emotion*group + (1|id), behav_by_run %>% filter(rewFunc %in% c("IEV", "DEV"))))
car::Anova(m6, type=3)
plot_model(m6, type="pred", terms=c("group", "rewFunc", "emotion"))
```

Also no evidence that average choice quality is worse in BPD

Any evidence of age-related effects?

```{r}
summary(m7 <- lmer(earnings ~ run + rewFunc*emotion*BPQ_total*Age_c + (1|id), behav_by_run))
car::Anova(m7, type=3)
plot_model(m7, type="pred", terms=c("BPQ_total", "rewFunc", "emotion"))

summary(m8 <- lmer(earnings ~ run + rewFunc*emotion*BPQ_total*Age_c + (1|id), behav_by_run %>% filter(rewFunc %in% c("IEV", "DEV"))))
car::Anova(m8, type=3)
plot_model(m8, type="pred", terms=c("BPQ_total", "rewFunc", "emotion"))

summary(m9 <- lmer(earnings ~ run + rewFunc*emotion*group*Age_c + (1|id), behav_by_run %>% filter(rewFunc %in% c("IEV", "DEV"))))
car::Anova(m9, type=3)
plot_model(m9, type="pred", terms=c("group", "rewFunc", "emotion"))
```

Absolutely no evidence of age-related differences


## descriptive visualizations

```{r}
ggplot(behav, aes(x=run_trial, y=entropy, color=group)) + facet_grid(emotion~rewFunc) + stat_smooth(method="loess")
```

## entropy curvature differences

No evidence of group differences in overall entropy in the run
```{r}
summary(mmm <- lmer(entropy ~ emotion*rewFunc*group + (1 | id/run), 
                    behav %>% filter(rewFunc %in% c("IEV", "DEV")), REML=FALSE))
car::Anova(mmm)

```

What if we account for trial? It has a massive effect, but not obvious relationship to group

```{r}
behav$i_trial <- -100/behav$asc_trial
summary(mmm <- lmer(entropy ~ emotion*rewFunc*group + i_trial*group + (1 | id/run), 
                    behav %>% filter(rewFunc %in% c("IEV", "DEV")), REML=FALSE))
car::Anova(mmm)

```

## performance

What about basic RT sticking?


```{r}
summary(mmm <- lmer(rt_csv ~ (rt_lag + i_trial + BPQ_total + emotion + omission_lag + rewFunc)^3 + (1 | id/run), 
                    behav %>% filter(rewFunc %in% c("IEV", "DEV")), REML=FALSE))
car::Anova(mmm, type=3)
```

rew_om x group x rt_lag?

Looks like using BPQ_total gives a stronger signal and basically shows an impatience effect: those with BPD are less willing to wait after a reward.

```{r}
summary(mmm <- lmer(rt_csv ~ i_trial + rt_lag * BPQ_total * omission_lag + emotion + rewFunc + (1 | id/run), 
                    behav %>% filter(rewFunc %in% c("IEV", "DEV")), REML=FALSE))
car::Anova(mmm, type=3)

summary(mmm <- lmer(rt_csv ~ i_trial + rt_lag * BPQ_total * omission_lag + emotion + rewFunc + (1 | id/run), 
                    behav, REML=FALSE))
car::Anova(mmm, type=3)

#rt_lag:BPQ_total:omission_lag
plot_model(mmm, type="pred", terms=c("rt_lag", "BPQ_total", "omission_lag"))

```

```{r}
summary(mmm <- lmer(rt_csv ~ i_trial + (rt_lag + BPQ_total + omission_lag + emotion + rewFunc)^3 + (1 | id/run), 
                    behav %>% filter(rewFunc %in% c("IEV", "DEV")), REML=FALSE))
car::Anova(mmm, type=3)

summary(mmm <- lmer(rt_csv ~ i_trial + rt_lag*BPQ_total*emotion + omission_lag + rewFunc + (1 | id/run), 
                    behav %>% filter(rewFunc %in% c("IEV", "DEV")), REML=FALSE))

plot_model(mmm, type = "std")#, terms=c("rt_lag", "BPQ_total", "emotion"))
emtrends(mmm, ~emotion, var = "BPQ_total", data=behav)
```

## model-based analyses

```{r}
mmm <- lmer(rt_csv ~ (scale(-1/run_trial) + scale(rt_lag) + scale(rt_vmax_lag) + vmax_lag + rt_vmax_change + entropy + BPQ_total)^3 + (1|id/run), behav)
summary(mmm)
car::Anova(mmm, type=3)
```

